A comparison on R tools for nonlinear least squares modeling

### 2023-03-14 RefactoringNLS/ComparingNLStools.Rmd
update with citations and descriptions

Other CRAN packages worth Commentary: ??

??put in proper refs to each

onls (@R-onls-manual) -- when there are errors in variables, this minimizes the orthogonal 
residuals

crsnls -- Controlled random search. No vignette!

NISTnls -- NIST problems for testing nls methods

nlshelper -- A few utilities for summarizing, testing, and plotting non-linear
    regression models fit with nls(), nlsList() or nlme()

nlsic -- We solve non linear least squares problems with optional
    equality and/or inequality constraints.

nlsMicrobio -- Data sets and nonlinear regression models dedicated to predictive microbiology.

@@

nls.multstart -- Non-linear least squares regression with the Levenberg-Marquardt algorithm using multiple starting values for increasing the chance that the minimum found is the global minimum.

nlstac -- Set of functions implementing the algorithm described in Fernandez 
    Torvisco et al. (2018) for fitting separable nonlinear regression curves.
    ?? check ref and add to bib.??

nlstools -- Several tools for assessing the quality of fit of a gaussian nonlinear model are provided.

>>>>

There are a number of other CRAN packages worth commentary, and we will give
some capsule comments later. Packages outside of CRAN, even if they might be
worthy of review, will not be included here.

??move down
## Other CRAN packages

`onls` (@R-onls-manual) -- The usual optimization in estimating nonlinear models
is the vertical difference between the dependent variable and the functional
model. `onls` minimizes the sum of orthogonal residuals. The objective is
therefore different, and involves nontrivial extra calculation. A vignette with
the package and the blog article https://www.r-bloggers.com/2015/01/introducing-orthogonal-nonlinear-least-squares-regression-in-r/ 
give some description with illustrative graphs. `onls` appears to be limited to 
problems with one independent and one dependent variable. The Wikipedia article
https://en.wikipedia.org/wiki/Total_least_squares presents an overview of some
ideas. Our view is that the approach needs wider discussion and tutorial examples
to allow its merits to be judged.

`crsnls` (@R-crsnls-manual) -- This package allows nonlinear estimation by 
controlled random search via two methods. There is unfortunately no vignette.
A modest trial we carried out showed `nlsr::nlxb()` gave the same results in 
a small fraction of the time required by either of the methods in `crsnls`.

NISTnls (@NISTnls-manual) -- This package provides R code and data for a set of
nonlinear least squares problems from the U.S. National Institute for Standards
and Technology for testing solvers. We caution that the problems are generally 
poorly conditioned and may not represent real-world situations.

nlshelper (@nlshelper-manual) -- This package, which unfortunately lacks a
vignette, provides a few utilities for summarizing, testing, and plotting 
non-linear regression models estimated with nls(), nlsList() or nlme() that
are linked or grouped in some way.

nlsic (@nlsic-manual) -- This solves non linear least squares problems with optional
equality and/or inequality constraints. It is clearly not about modeling, and
the input and output are quite different from class `nls` methods.

`nlsMicrobio` (@nlsMicrobio-manual) -- Data sets and nonlinear regression models dedicated to predictive microbiology, including a vignette, by authors of the
`nlstools` package who have a strong record in the subject.

`nlstools` (@p-nlstools) -- Several tools for assessing the quality of fit of a gaussian nonlinear model are provided.

@@

`nls.multstart` (@nlsmultstart) -- Non-linear least squares regression using
AIC scores with the Levenberg-Marquardt algorithm using multiple starting values for increasing the chance that the minimum found is the global minimum.

`nls2` (@nls2-22) -- Nonlinear least squares by brute force has some similar
motivations, but a longer history and using `nls()` within multiple trials.

`nlstac` (@m-nlstac) -- Set of functions implementing the algorithm described in Fernandez 
    Torvisco et al. (2018) for fitting separable nonlinear regression curves.

    ?? check ref and add to bib.??
